---
title: "CellProfiler Output Cleaning"
author: "gianna"
date: "2026-06-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyr)
library(tidyverse)
library(dplyr)
library(stringr)

setwd("~/Documents/GitHub/chRonic-time-wasting-disease/CellProfilerOutputs/CellProfilerData")

```
```{r}
#join together the separate files
Image_Control<-read.csv("~/Documents/GitHub/chRonic-time-wasting-disease/CellProfilerOutputs/CellProfilerData/pRions_v3_Control_Image.csv")
Image_Deer<-read.csv("~/Documents/GitHub/chRonic-time-wasting-disease/CellProfilerOutputs/CellProfilerData/pRions_v3_GTDeer_Image.csv")
Image_Elk<-read.csv("~/Documents/GitHub/chRonic-time-wasting-disease/CellProfilerOutputs/CellProfilerData/pRions_v3_GTElk_Image.csv")
pRions_Image<- rbind(Image_Control, Image_Deer, Image_Elk)


Depo_Control<-read.csv("~/Documents/GitHub/chRonic-time-wasting-disease/CellProfilerOutputs/CellProfilerData/pRions_v3_Control_DepositionsinTissue.csv")
Depo_Deer<-read.csv("~/Documents/GitHub/chRonic-time-wasting-disease/CellProfilerOutputs/CellProfilerData/pRions_v3_GTDeer_DepositionsinTissue.csv")
Depo_Elk<-read.csv("~/Documents/GitHub/chRonic-time-wasting-disease/CellProfilerOutputs/CellProfilerData/pRions_v3_GTElk_DepositionsinTissue.csv")
pRions_v3_DepTissue<- rbind(Depo_Control, Depo_Deer, Depo_Elk)


Tissue_Control<-read.csv("~/Documents/GitHub/chRonic-time-wasting-disease/CellProfilerOutputs/CellProfilerData/pRions_v3_Control_ShrunkenTissue2.csv")
Tissue_Deer<-read.csv("~/Documents/GitHub/chRonic-time-wasting-disease/CellProfilerOutputs/CellProfilerData/pRions_v3_GTDeer_ShrunkenTissue2.csv")
Tissue_Elk<-read.csv("~/Documents/GitHub/chRonic-time-wasting-disease/CellProfilerOutputs/CellProfilerData/pRions_v3_GTElk_ShrunkenTissue2.csv")

#tissue area/shape measure can be used for later downstream analyses that can give info on the health of the brain (perimeter-area ratios, eccentricity values, etc) but outside the scope of what we aimed for today

```

## pRions_v3_Depositions
First, we would like to compare the overall image data to determine differences in deposition counts and mean area of depositions
```{r ImagesFile}

pRions_Image <- pRions_Image%>%
  mutate(FileName = FileName_DNA)  %>%
  separate(col = FileName_DNA, 
           into = c("Mouse", "Treatment", "Region", "Magnification", "ImageNum"), 
           sep = "_", fill = "right", extra = "drop") %>% 
  select("Mouse", "Treatment", "Region", "Magnification", "ImageNum", "Count_DepositionsinTissue", "Mean_DepositionsinTissue_AreaShape_Area", "AreaOccupied_AreaOccupied_DepositionsinTissue","AreaOccupied_AreaOccupied_ShrunkenTissue2") %>%
  rename(Count_Depo = Count_DepositionsinTissue) %>%
  rename(Mean_Depo_Area = Mean_DepositionsinTissue_AreaShape_Area) %>%
  rename(DepoAreaOccupied = AreaOccupied_AreaOccupied_DepositionsinTissue) %>%
  rename(TissueArea = AreaOccupied_AreaOccupied_ShrunkenTissue2) %>%
  mutate(PercentOccupied = DepoAreaOccupied/TissueArea*100)
  

```

## plotting image-level data:

Counts Plots
```{r CountsImage}
library(tidyverse)
library(ggprism)
library(ggsignif)

ggplot(
  pRions_Image,
  aes(
    x = Mouse,
    y = Count_Depo,
    fill = Mouse
  )
) +
  geom_boxplot(
    outlier.shape = NA,
    alpha = 0.8
  ) +
  geom_jitter(
    width = 0.15,
    alpha = 0.6,
    size = 2
  ) +
  facet_wrap(~Region) +
  theme_prism(base_size = 14) +
  labs(
    x = "Mouse",
    y = "Depositions per Image"
  ) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(
      angle = 45,
      hjust = 1
    )
  )+
  scale_y_continuous(
  expand = expansion(mult = c(0.05, 0.2))
) +
  geom_signif(
  comparisons = list(
    c("GtDeer", "WT"), 
    c("GtElk", "WT"), 
    c("GtDeer", "GtElk")
  ),
  map_signif_level = TRUE,
  test = "t.test")

```
And % area occupied:
```{r AreaPercent}
ggplot(
  pRions_Image,
  aes(
    x = Mouse,
    y = PercentOccupied,
    fill = Mouse
  )
) +
  geom_boxplot(
    outlier.shape = NA,
    alpha = 0.8
  ) +
  geom_jitter(
    width = 0.15,
    alpha = 0.6,
    size = 2
  ) +
  facet_wrap(~Region) +
  theme_prism(base_size = 14) +
  labs(
    x = "Mouse",
    y = "Percent Area Occupied Depositions"
  ) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(
      angle = 45,
      hjust = 1
    )
  )+
  scale_y_continuous(
  expand = expansion(mult = c(0.05, 0.2))
) +
  geom_signif(
  comparisons = list(
    c("GtDeer", "WT"), 
    c("GtElk", "WT"), 
    c("GtDeer", "GtElk")
  ),
  map_signif_level = TRUE,
  test = "t.test")

```


# Next, we would like to look at the properties of each individual deposition

WT/Control does not have significant numbers of depositions (just random artifacts) and was not included in this comparison

Clean Deposition Data
```{r DepositionClean}


pRion_DepTissue <- pRions_v3_DepTissue %>%
  mutate(FileName = FileName_DNA)  %>%
  separate(col = FileName_DNA, 
           into = c("Mouse", "Treatment", "Region", "Magnification", "ImageNum"), 
           sep = "_", fill = "right", extra = "drop") %>%
  rename(Area = AreaShape_Area) %>% 
  select("Mouse", "Treatment", "Region", "Magnification", "ImageNum", "Area", "Intensity_IntegratedIntensity_InvertedColor", "Intensity_MaxIntensity_InvertedColor", "Intensity_MinIntensity_InvertedColor") %>% #integrated intensity represents area x mean intensity
  rename(Intensity = Intensity_IntegratedIntensity_InvertedColor) %>%
  rename(IntensityMax = Intensity_MaxIntensity_InvertedColor) %>%
  rename(IntensityMin = Intensity_MinIntensity_InvertedColor) %>% 
  filter(Treatment != "control")


```
Plotting each deposition for area:
```{r Area plot}

ggplot(
  pRion_DepTissue,
  aes(
    x = Mouse,
    y = Area,
    fill = Mouse
  )
) +
  geom_boxplot(
    width = 0.7,
    outlier.shape = NA,
    alpha = 0.6,
    position = position_dodge(width = 0.8)
  ) +
  geom_point(
    aes(color = Mouse),
    position = position_jitterdodge(
      jitter.width = 0.15,
      dodge.width = 0.8
    ),
    size = 2,
    alpha = 0.005
  ) +
  scale_y_log10(
  expand = expansion(mult = c(0.05, 0.15))
) +
  theme_prism(base_size = 14) +
  labs(
  title = "Prion Deposition Area by Mouse Strain and Treatment",
  x = "Mouse Strain",
  y = expression("Deposition Area (pixels"^2*")"),
  fill = "Treatment",
  color = "Treatment"
  ) +
  theme(
  plot.title = element_text(
    hjust = 0.5,
    size = 12,
    face = "bold"
  ),
  axis.text.x = element_text(
    angle = 45,
    hjust = 1
  )
)+
  geom_signif(
    comparisons = list(c("GtDeer", "GtElk")), 
    map_signif_level = TRUE,                 
    test = "t.test"                         
  )
```
Now looking at intensity:

```{r Integrated Intensity Plot}
ggplot(
  pRion_DepTissue,
  aes(
    x = Mouse,
    y = Intensity,
    fill = Mouse
  )
) +
  geom_boxplot(
    width = 0.7,
    outlier.shape = NA,
    alpha = 0.6,
    position = position_dodge(width = 0.8)
  ) +
  geom_point(
    aes(color = Mouse),
    position = position_jitterdodge(
      jitter.width = 0.15,
      dodge.width = 0.8
    ),
    size = 2,
    alpha = 0.005
  ) +
  scale_y_log10(
  breaks = c(10, 100, 1000, 10000),
  labels = scales::label_number(),
  limits = c(10, 10000),
  expand = expansion(mult = c(0.05, 0.15))
) +
  theme_prism(base_size = 14) +
  labs(
    title = "Prion Deposition Staining Intensity by Mouse Strain and Treatment",
    x = "Mouse Strain",
    y = "Integrated Intensity",
    fill = "Treatment",
    color = "Treatment"
  ) +
  theme(
  plot.title = element_text(
    hjust = 0.5,
    size = 12,
    face = "bold"
  ),
  axis.text.x = element_text(
    angle = 45,
    hjust = 1
  )
)+
  geom_signif(
    comparisons = list(c("GtDeer", "GtElk")), 
    map_signif_level = TRUE,                 
    test = "t.test"                         
  )
```







